---
output: html_document
---
# Fiji islands earthquake data analysis
## data consisting of 1000 earthquake documentations we are analyzing earthquake magnitude distribution, specific dangerous areas and more.

####loading required packages
```{r setup, include=FALSE}
library("ggplot2")
library("ggmap")
library("caret")
```
library("ggplot2")
library("ggmap")
library("caret")

####loading life 
```{r read csv, include=FALSE}
df = read.csv("c:\\dev\\data\\quakes.csv",TRUE)
```
df = read.csv("c:\\dev\\data\\quakes.csv",TRUE)
#### showing data format and the few first rows 
```{r head, include=FALSE}
head(df)
```
head(df)
#### showing summary
```{r summary, include=FALSE}
summary(df)
```
summary(df)
#### showing magnitude distribution
```{r boxplot, include=FALSE}
boxplot(df$mag,col="red")
```
boxplot(df$mag,col="red")
####create magnitude histogram to visualize frequency of different magnitude quakes
#### red line shows median
#### yellow line shows mean
```{r hist mag, include=FALSE}
hist(df$mag,col="blue")
abline(v=median(df$mag),lwd=4,col="red")
abline(v=mean(df$mag),lwd=4,col="yellow")
```
hist(df$mag,col="blue")
abline(v=median(df$mag),lwd=4,col="red")
abline(v=mean(df$mag),lwd=4,col="yellow")
####scatter plot of quakes to visualize dangerous areas by latitude
```{r scatter lat, include=FALSE}
with(df,plot(lat,mag))
```
with(df,plot(lat,mag))
####scatter plot of quakes to visualize dangerous areas by longtitude
```{r scatter_long, include=FALSE}
with(df,plot(mag,long))
```
####scatter plot of both latitude and longtitude. magnitude is shown by color axis.
```{r scatter_both, include=FALSE}
qplot(long,lat,data=df,col=mag)
map = get_map (location = c(lon=mean(df$long),lat=mean(df$lat)),zoom=5,maptype="satellite",scale=2)
```
qplot(long,lat,data=df,col=mag)
map = get_map (location = c(lon=mean(df$long),lat=mean(df$lat)),zoom=5,maptype="satellite",scale=2)
####showing quake distibution on map
```{r map1, include=FALSE}
ggmap(map) + geom_point(data = df, aes(x = long, y = lat), size = 1,color="red") +
  guides(fill=FALSE, alpha=FALSE, size=FALSE)
```
ggmap(map) + geom_point(data = df, aes(x = long, y = lat), size = 1,color="red") +
  guides(fill=FALSE, alpha=FALSE, size=FALSE)
####showing the most dangerous places on fiji
```{r map2, include=FALSE}
ggmap(map) + geom_point(data = df[df$mag>5.8,], aes(x = long, y = lat), size = 2,color="red") +
  guides(fill=FALSE, alpha=FALSE, size=FALSE)

```
ggmap(map) + geom_point(data = df[df$mag>5.8,], aes(x = long, y = lat), size = 2,color="red") +
  guides(fill=FALSE, alpha=FALSE, size=FALSE)

####check for correlation between depth and magnitude
```{r corel, include=FALSE}
cordf = cor(df[,c("mag","stations")])
cordf[2,1]
```
cordf = cor(df[,c("mag","stations")])
cordf[2,1]
#### we can see there is a high correlation between number of stations detecting the quake and its magnitude.
#### thus, we can assume that this feature is redundant.
#### moreover, high correlation also implies that one parameter can predict the other.
#### This means that if a sesmologic noise is caught on many stations it could predict a quake of big magnitude.
#### this can also be shown be a line graph
```{r lineg, include=FALSE}
ggplot(data=df,aes(x=mag,y=stations,group=1))+geom_line(data=df,stat="identity")

```
ggplot(data=df,aes(x=mag,y=stations,group=1))+geom_line(data=df,stat="identity")